### 2024-05-05 Papers/Graph Algorithms/Booklet/GA.tex
initial Commit for create GA.tex and GA.pdf

\documentclass[a4paper,10pt]{book}

\usepackage{geometry}
\geometry{
    right=1cm,
    left=1cm,
    top=2cm,
    bottom=2cm
}

\usepackage{xepersian}
\settextfont{Vazirmatn-Regular.ttf}

\title{\Huge جزوه الگوریتم‌های گراف}
\author{استاد مربوطه:\\سرکار خانم دکتر معصومه دامرودی}
\date{نویسنده:\\محمد خورشیدی روزبهانی}

\begin{document}

    \maketitle

    سلام

    \tableofcontents

    \chapter{مقدمات}

        سلام

\end{document}