### 2024-05-08 Papers/Graph Algorithms/Exam/Mid-Term-1.tex
init Mid-Term-1

\documentclass[a4paper,10pt]{article}

\begin{document}

    hello

\end{document}