### 2024-06-29 Papers/Cloud Computing/Presentation-Article/Article.tex
Commit

\documentclass[a4paper,10pt]{article}

\usepackage{geometry}
\geometry{
    a4paper,
    left=1cm,
    right=1cm,
    top=1cm,
    bottom=2cm}

\usepackage{xepersian}
\settextfont{Vazirmatn-Regular.ttf}

\title{عنوان}
\author{محمد خورشیدی روزبهانی\and 40215741002013}
\date{}

\linespread{1.5}

\begin{document}

    \maketitle

    \begin{abstract}

        چکیده

    \end{abstract}

\end{document}