### 2024-09-11 Papers/Parallel Aalgorithms/Booklet/PA.tex
Commit

\documentclass[a4paper, 10pt]{book}

\usepackage{fontspec} % --fontspec-- package to use other fonts
\setmainfont{Vazirmatn-Regular.ttf} 
\newfontfamily\bold{Vazirmatn-Bold.ttf}

\usepackage{geometry}
\geometry{top=2cm, bottom=2cm, left=1.5cm, right=1.5cm}

\title{\huge {\bold Parallel Algorithms}}
\author{Professor\\{\bold Dr. Masoumeh Damroudi}}
\date{Author\\{\bold Mohammad Khorshidi}}

\begin{document}

    \pagenumbering{roman}
    \maketitle
    \newpage

    \begin{center}
        \Large {\bold Course Overview}
    \end{center}

    \section{Course Structure}

    \section{Instructor's Approach}

    \section{Grading and Evaluation}

    \section{Course Materials and Resources}

    \tableofcontents

    \chapter{{\bold Introduction}}

    \pagenumbering{arabic}

        some text ...

\end{document}