### 2024-09-18 Papers/Parallel Aalgorithms/Presentation/Article-fa.tex
Commit

\documentclass[a4paper,10pt]{article}

\usepackage{geometry}
\geometry{top=2cm, bottom=2cm, left=1.5cm, right=1.5cm}

\usepackage{xepersian}
\settextfont{Vazirmatn-Regular.ttf}

\title{\huge بهینه‌سازی کلونی مورچه‌ها به‌صورت موازی بر روی پردازنده‌های چند هسته‌ای SIMD}
\author{محمد خورشیدی روزبهانی\\40215741002013 \and شارا شاهوردیان\\40215741002032}
\date{}

\linespread{1.5}

\begin{document}

    \maketitle

    % \vspace{5pt}

    \begin{abstract}

        متن

    \end{abstract}

    % \vspace{15pt}

    \section{مقدمه}

        متن

\end{document}