### 2024-10-13 Papers/Parallel Aalgorithms/Practice/Practice 1.tex
Initialize Practice 1.tex
- Use --fancyhdr-- package for header and footer

\documentclass[a4paper,10pt]{article}

\usepackage{geometry}
\geometry{top=2cm, bottom=2cm, left=1.5cm, right=1.5cm}

\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
\fancyhead[R]{\today}
\fancyhead[C]{Practice 1}
\fancyhead[L]{Parallel Algorithms}
\fancyfoot[R]{\thepage}
\fancyfoot[L]{M. Khorshidi R. - 40215741002013}

\usepackage{fontspec}
\setmainfont{Vazirmatn-Regular.ttf}
\newfontfamily\bold{Vazirmatn-Bold.ttf}

\linespread{1.3}

\begin{document}

    \paragraph{\bold Q:} Some text ...

    \paragraph{\bold A:} Some text ...

    \noindent
    \hrulefill

    \paragraph{\bold Q:} Some text ...

    \paragraph{\bold A:} Some text ...

\end{document}